Hydrodynamics of two vesicles interacting via a Van der Waals potential

### 2018-12-29 vesicle_code/docs/adhesion/adhesion.tex
removed yuan's reference file

When two vesicles move towards each other under a constant force $F$ without any external fluid flow, the height  $(h)$ of the lubrication film between two vesicles follows the
draining dynamics \cite{RamachandranLeal2010_PoF}
\begin{equation}
\label{eq:dhdt}
\frac{d h}{dt} \sim \frac{K_a^{2/3} F^{1/3}}{\mu R_0^{10/3}} h^3,
\end{equation}
where $K_a$ is the area expansion modulus, $\mu$ is the viscosity of exterior fluid, $R_0$ is the radius of the undeformed vesicle.
When $F$ is a constant (independent of separation distance $h$ and/or time $t$), the above scaling can give variation of $h$ versus $t$ by integrating the above equation
\begin{equation}
t \sim \frac{\mu R_0^{10/3}}{K_a^{2/3} F^{1/3}}\frac{1}{2 h^2} \bigg|^{h(t)}_{h(0)},
\end{equation}
where $h(0)$ is the vesicle separation at $t=0$.
We note that $F< (>) 0$ for an attractive  (repulsive) force and consequently $h(t)$ decreases (increases) from  the initial separation distance $h(0)$. 
When the force is attractive, $h$ decreases monotonically and $t$ in the above
equation is the
draining time that diverges as $h\rightarrow 0$.
%Similarly $F>0$ for a repulsive force and consequently $h(t) > h(0)$.

When the force on each vesicle is a function of separation distance $h$ of the form (with an integer $m\ge 2$)
\begin{equation}
\label{eq:Fh}
F(h) \sim \sigma \frac{\delta^m}{h^{m+1}}\left(\left(\frac{\delta}{h}\right)^2-1\right),
\end{equation}
attraction is dominant at ``large" scales ($h > \delta$) while repulsion is dominant at ``small" scales ($h<\delta$).
Integration of equation~\ref{eq:dhdt} with the above $F(h)$ gives the relationship between $t$ and $h$.
In this case the relationship involves an integral of a function of dimensionless $\bar{h} \equiv h/\delta$:
\begin{align}
t\frac{K_a^{2/3}}{\mu R_0^{10/3}}& = \int^{h(t)}_{h(0)}-\frac{dh}{h^3 F^{1/3}} = -\frac{1}{\sigma^{1/3}\delta^{5/3}}\int^{\bar{h}(t)}_{\bar{h}(0)}\frac{d \bar{h}}{\bar{h}^{(8-m)/3}(\bar{h}^2-1)^{1/3}},
\end{align}
\begin{equation}
\label{eq:teq_scaling}
t\sim \frac{\mu R_0^{10/3}}{K_a^{2/3}}\frac{1}{\sigma^{1/3}\delta^{5/3}}.
\end{equation}
Equation~\ref{eq:teq_scaling} tells us that both adhesion strength $\sigma$ and the separation distance $\delta$ affect
the time it takes for a pair of vesicles to reach equilibrium under adhesion force $F$ in equation~\ref{eq:Fh}.  In particular, the equilibration time $t$ 
is proportional to both (1) the separation distance $\delta^{-5/3}$ (thus $t\rightarrow \infty$ as $\delta\rightarrow 0$), and (2) the adhesion strength $\sigma^{-1/3}$ 
(as in constant forcing case \cite{RamachandranLeal2010_PoF}).  From the above analysis we also expect that the scaling of $t$ with respect to $\delta$ depends on the
adhesion potential, while the scaling with respect to adhesion strength is independent of the specific form of the potential.

@@

\subsection{Effects of the adhesion parameters on equilibrium}
\label{subsec:qflow_adhesion_parameters} 
Here we investigate the dynamics and steady state configuration of two
identical vesicles placed next to each other in a quiescent flow.  For a
large reduced area, the two vesicles move toward each other until the
membrane flattens in the adhesion region, as shown in
Figures~\ref{fig:qflow01} and~\ref{fig:qflow02}.  The color coding along
the membrane denotes the membrane tension, which is the lowest in the
adhesion region.

\begin{figure}
\includegraphics[keepaspectratio=true,scale=0.125]{figs/Nrelax2Ves02-rA0p96adR0p4.png}
\caption{Two vesicles interacting with each other via a vdW potential at different strengths in a quiescent viscous fluid. The color coding is the tension along the vesicle.}
\label{fig:qflow01}
\end{figure}

\begin{figure}
\includegraphics[keepaspectratio=true,scale=0.25]{figs/Nrelax2Ves02composite-draining.pdf}
\caption{Adhesion of two vesicles in a quiescent flow.  Only one vesicle
  is illustrated, and the second is a mirror image.  Left panel: The
  steady state configuration for different vdW interaction distances.
  The color coding is the tension along the vesicle.  Right panel:
  Vesicle configurations at different times for different vdW strengths.
  The top plot has the weakest vdW strength and the bottom plot has the
  strongest vdW strength.}
\label{fig:qflow02} 
\end{figure}

% \begin{figure}
% \includegraphics[keepaspectratio=true,scale=0.4]{figs/relax2Ves01g-draining01a.eps}
% \caption{Adhesion dynamics of two vesicles with $\delta = 0.3$. From $t=0.3$ to $t=40$, the time interval $\delta t = 0.3$ between every two neighboring dash-dotted curves.}
% \label{fig:qflow_g-draining01a}
% \end{figure}

% \begin{figure}
% \includegraphics[keepaspectratio=true,scale=0.4]{figs/relax2Ves01j-draining01a.eps}
% \caption{Adhesion dynamics of two vesicles with $\delta = 0.6$. From $t=0.1$ to $t=40$, the time interval $\delta t = 0.1$ between every two neighboring dash-dotted curves.}
% \label{fig:qflow_g-draining01a}
% \end{figure}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsection{Effects of reduced area}
\label{sec:qflow_reduced_area} 
Figure~\ref{fig:qflow_reduced_area}

\begin{figure}
\includegraphics[keepaspectratio=true,scale=0.5]{figs/nrelax2Ves02o_rA.png}
\caption{The effect of the reduced area on adhering vesicles in a
quiescent flow.  The parameters of the vdW potential are $\delta = 0.2$
and $\sigma = 0.5$.}
\label{fig:qflow_reduced_area}
\end{figure}

% \begin{figure}
% \includegraphics[keepaspectratio=true,scale=0.4]{figs/relax2Ves01usa.eps}
% \caption{Adhesion dynamics of two vesicles with $\delta = 0.4$ and $\sigma=0.2$. From $t=0.06$ to $t=100$, the time interval $\delta t = 0.06$ between every two neighboring dash-dotted curves.}
% \label{fig:qflow_u-draining}
% \end{figure}

% \begin{figure}
% \includegraphics[keepaspectratio=true,scale=0.4]{figs/relax2Ves01zsa.eps}
% \caption{Adhesion dynamics of two vesicles with $\delta = 0.4$ and $\sigma=0.2$. From $t=1$ to $t=100$, the time interval $\delta t = 1$ 
% between every two neighboring dash-dotted curves.}
% \label{fig:qflow_z-draining}
% \end{figure}

@@

\bibliography{refs,refs_yny}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%\subsection{\label{sec:level2}Second-level heading: Formatting}
%
%This file may be formatted in either the \texttt{preprint} or
%\texttt{reprint} style. \texttt{reprint} format mimics final journal output. 
%Either format may be used for submission purposes. \texttt{letter} sized paper should
%be used when submitting to APS journals.
%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%\subsubsection{Wide text (A level-3 head)}
%The \texttt{widetext} environment will make the text the width of the
%full page, as on page~\pageref{eq:wideeq}. (Note the use the
%\verb+\pageref{#1}+ command to refer to the page number.) 
%\paragraph{Note (Fourth-level head is run in)}
%The width-changing commands only take effect in two-column formatting. 
%There is no effect if text is in a single column.

>>>>

When two vesicles move towards (or away from) each other under a constant force $F$ without any external fluid flow, the height  $(h)$ of the thin liquid film between two vesicles follows the
draining dynamics \cite{RamachandranLeal2010_PoF}
\begin{equation}
\label{eq:dhdt}
\frac{d h}{dt} \sim \frac{K_a^{2/3} F^{1/3}}{\mu R_0^{10/3}} h^3,
\end{equation}
where $K_a$ is the area expansion modulus, $\mu$ is the viscosity of exterior fluid, $R_0$ is the radius of the undeformed vesicle.
For the case of a constant forcing $F$ (independent of separation distance $h$ and time $t$), equation~\ref{eq:dhdt} can be easily integrated to give
a relation between film thickness $h$ and time $t$:
\begin{equation}
t \sim \frac{\mu R_0^{10/3}}{K_a^{2/3} F^{1/3}}\frac{1}{2 h^2} \bigg|^{h(t)}_{h(0)},
\end{equation}
where $h(0)$ is the vesicle separation at $t=0$.
We note that $F< (>) 0$ for an attractive  (repulsive) interaction between two vesicles,  and consequently $h(t)$ decreases (increases) from  the initial separation distance $h(0)$. 
When the force is attractive, $h$ decreases monotonically and $t$ in the above
equation is the
draining time that diverges as $h\rightarrow 0$.
%Similarly $F>0$ for a repulsive force and consequently $h(t) > h(0)$.

When the force on each vesicle is a function of separation distance $h$ of the form (with an integer $m\ge 2$)
\begin{equation}
\label{eq:Fh}
F(h) \sim \sigma \frac{\delta^m}{h^{m+1}}\left(\left(\frac{\delta}{h}\right)^2-1\right),
\end{equation}
attraction is dominant at ``large" scales ($h > \delta$) while repulsion is dominant at ``small" scales ($h<\delta$).
Integration of equation~\ref{eq:dhdt} with the above $F(h)$ gives the relationship between $t$ and $h$.
In this case the relationship involves an integral of a function of dimensionless $\bar{h} \equiv h/\delta$:
\begin{align}
t\frac{K_a^{2/3}}{\mu R_0^{10/3}}& = \int^{h(t)}_{h(0)}-\frac{dh}{h^3 F^{1/3}} = -\frac{1}{\sigma^{1/3}\delta^{5/3}}\int^{\bar{h}(t)}_{\bar{h}(0)}\frac{d \bar{h}}{\bar{h}^{(8-m)/3}(\bar{h}^2-1)^{1/3}},
\end{align}
\begin{equation}
\label{eq:teq_scaling}
t\sim \frac{\mu R_0^{10/3}}{K_a^{2/3}}\frac{1}{\sigma^{1/3}\delta^{5/3}}.
\end{equation}
Equation~\ref{eq:teq_scaling} tells us that both adhesion strength $\sigma$ and the separation distance $\delta$ affect
the time it takes for a pair of vesicles to reach equilibrium under adhesion force $F$ in equation~\ref{eq:Fh}.  In particular, the equilibration time $t$ 
is proportional to (1) the separation distance $\delta^{-5/3}$ (thus $t\rightarrow \infty$ as $\delta\rightarrow 0$), and (2) the adhesion strength $\sigma^{-1/3}$ 
(as in constant forcing case \cite{RamachandranLeal2010_PoF}).  From the above analysis we also expect that the scaling of $t$ with respect to $\delta$ depends on the
adhesion potential, while the scaling with respect to adhesion strength is independent of the specific form of the potential.

@@

\subsection{Effects of the adhesion parameters on equilibrium configuration of a vesicle pair}
\label{subsec:qflow_adhesion_parameters} 
In the simulations of a pair of 2d vesicles interacting with each other via the adhesion force in equation~\ref{eq:Fh} 
two vesicles are placed (not too far apart for numerical convenience) in a quiescent flow. Long-range attraction pulls two vesicles together as they deform until their
separation distance is close to $\delta$ (where $F$ vanishes). 
\begin{figure}
\includegraphics[keepaspectratio=true,scale=0.25]{figs/Dec18_vesicle_shape_vs_rA_00.jpeg}
\caption{Equilibrium shapes of wo vesicles under adhesive  interactions with adhesion strength $\sigma=0.5$ and separation distance $\delta=0.2$. 
Vesicle reduced area $\Delta A=0.95$, $0.75$ and $0.6$ as labeled.
 The color coding is the tension along the vesicle.}
\label{fig:Dec18_vesicle_shape}
\end{figure}

Figure~\ref{fig:Dec18_vesicle_shape}(a) shows the equilibrium shapes of vesicles with
three values of reduced area: $\Delta A=0.95$, $0.75$ and $0.6$ with $\sigma=5$ and $\delta = 0.2$.
The equilibrium vesicle shape for $\Delta A=0.95$ is the circular cap with a flatten contact region. 
This is similar to the observed shapes of two vesicles under strong adhesive interaction in \cite{RamachandranAndersonLealIsraelachvili2010_Langmuir}.

When vesicles are more deflated with a reduced area  $\Delta A = 0.75$, the equilibrium vesicle shape is elongated with a bigger contact region.
As the reduced area decreases further, we observe undulation of the elongated equilibrium vesicle shape on the non-contact side while the contact region remains flat. 
The color coding along each curve is for the tension of vesicle membrane.
We observe that the membrane tension in the contact region is very negative, indicating a dominant compression of membrane when the adhesion force is strong to keep the vesicles bound together.

\begin{figure}
\includegraphics[keepaspectratio=true,scale=0.125]{figs/Dec18_Eb_vs_rA_adR0p2_adS502.jpeg}
\includegraphics[keepaspectratio=true,scale=0.125]{figs/Dec18_Eadh_vs_rA_adR0p2_adS502.jpeg}
\includegraphics[keepaspectratio=true,scale=0.125]{figs/Dec18_EbEadh_vs_rA_adR0p2_adS502.jpeg}
\caption{Total bending energy (panel (a)), adhesion energy (panel (b)) and the sum of two energies (panel (c)) plotted against reduced area $\Delta A$.}
\label{fig:Dec18_vesicle_equilibrium1}
\end{figure}
Figure~\ref{fig:Dec18_vesicle_equilibrium1}(a) shows the total bending energy as a function of reduced area. The smaller the reduced area, the more vesicle area is available
for deformation and thus the bending energy is higher. Figure~\ref{fig:Dec18_vesicle_equilibrium1}(b) shows the total adhesion energy as a function of reduced area. We note
that the total adhesion energy is more negative as reduced area decreases. The sum of two energies is plotted in figure~\ref{fig:Dec18_vesicle_equilibrium1}(c), where a local minimum in the total energy is
found around $\Delta A =0.8$.

\begin{figure}
\includegraphics[keepaspectratio=true,scale=0.18]{figs/Dec18_Eb_vs_sigma_rA0p9502.jpeg}
\includegraphics[keepaspectratio=true,scale=0.18]{figs/Dec18_Eadh_vs_sigma_rA0p9502.jpeg}
\caption{Total bending energy (a) and adhesion energy (b) versus the adhesion strength $\sigma$ 
for two vesicles with reduced area $\Delta A=0.95$, separation distance $\delta = 0.2$ (triangles), $0.3$ (crosses)
and $0.4$ (circles). }
\label{fig:Dec18_equilibrium} 
\end{figure}
%
Figure~\ref{fig:Dec18_equilibrium} is an example of how the adhesion strength $\sigma$ and separation distance $\delta$ affect the equilibrium configuration of two vesicles under adhesive interactions.
For both vesicles the reduced area $\Delta A=0.95$ and the bending rigidity is $1$.
In figure~\ref{fig:Dec18_equilibrium}(a) 
the total bending energy at equilibrium is plotted against $\sigma$ for three values of separation distance $\delta$.
We observe that for $\Delta A=0.95$ the equilibrium vesicle shape does not vary much with the adhesion strength $\sigma$, while the total adhesion energy in figure~\ref{fig:Dec18_equilibrium}(b) varies linearly
with $\sigma$.

%% \begin{figure}
%% \includegraphics[keepaspectratio=true,scale=0.4]{figs/relax2Ves01g-draining01a.eps}
%% \caption{Adhesion dynamics of two vesicles with $\delta = 0.3$. From $t=0.3$ to $t=40$, the time interval $\delta t = 0.3$ between every two neighboring dash-dotted curves.}
%% \label{fig:qflow_g-draining01a}
%% \end{figure}
%
%% \begin{figure}
%% \includegraphics[keepaspectratio=true,scale=0.4]{figs/relax2Ves01j-draining01a.eps}
%% \caption{Adhesion dynamics of two vesicles with $\delta = 0.6$. From $t=0.1$ to $t=40$, the time interval $\delta t = 0.1$ between every two neighboring dash-dotted curves.}
%% \label{fig:qflow_g-draining01a}
%% \end{figure}
%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%\subsection{Effects of reduced area}
%\label{sec:qflow_reduced_area} 
%Figure~\ref{fig:qflow_reduced_area}
%
%\begin{figure}
%\includegraphics[keepaspectratio=true,scale=0.5]{figs/nrelax2Ves02o_rA.png}
%\caption{The effect of the reduced area on adhering vesicles in a
%quiescent flow.  The parameters of the vdW potential are $\delta = 0.2$
%and $\sigma = 0.5$.}
%\label{fig:qflow_reduced_area}
%\end{figure}
%
%% \begin{figure}
%% \includegraphics[keepaspectratio=true,scale=0.4]{figs/relax2Ves01usa.eps}
%% \caption{Adhesion dynamics of two vesicles with $\delta = 0.4$ and $\sigma=0.2$. From $t=0.06$ to $t=100$, the time interval $\delta t = 0.06$ between every two neighboring dash-dotted curves.}
%% \label{fig:qflow_u-draining}
%% \end{figure}
%
%% \begin{figure}
%% \includegraphics[keepaspectratio=true,scale=0.4]{figs/relax2Ves01zsa.eps}
%% \caption{Adhesion dynamics of two vesicles with $\delta = 0.4$ and $\sigma=0.2$. From $t=1$ to $t=100$, the time interval $\delta t = 1$ 
%% between every two neighboring dash-dotted curves.}
%% \label{fig:qflow_z-draining}
%% \end{figure}

@@

\bibliography{refs}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%\subsection{\label{sec:level2}Second-level heading: Formatting}
%
%This file may be formatted in either the \texttt{preprint} or
%\texttt{reprint} style. \texttt{reprint} format mimics final journal output. 
%Either format may be used for submission purposes. \texttt{letter} sized paper should
%be used when submitting to APS journals.
%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%\subsubsection{Wide text (A level-3 head)}
%The \texttt{widetext} environment will make the text the width of the
%full page, as on page~\pageref{eq:wideeq}. (Note the use the
%\verb+\pageref{#1}+ command to refer to the page number.) 
%\paragraph{Note (Fourth-level head is run in)}
%The width-changing commands only take effect in two-column formatting. 
%There is no effect if text is in a single column.
